---
title: "Bank Customer Churn Prediction Model"
author: "Alvin"
date: "2024-07-25"
output: pdf_document
---

## Objective:
The main aim of this is to predict the customer turn over of a bank from a sample dataset.

## Setting the working directory
```{r}
setwd("D:/R Programming")
getwd()
```
## Loading the required libraries
```{r}
library(caTools)
library(tidyverse)
library(dplyr)
library(corrplot)
library(ROCR)
```
## Loading the dataset
```{r}
df <- read.csv("C:/Users/Alvin/Downloads/Compressed/Bank Customer Churn/Customer-Churn-Records.csv")
```

## Exploring the dataset
## An overall summary of the dataset
```{r}
summary(df)
```
## Checking the first 5 rows of the dataset
```{r}
head(df)
```
## Cleaning the dataset
## Removing unwanted columns
```{r}
df1 <- df[-c(1, 2, 3, 5, 6, 17)]
```

## Checking for any outliers
```{r}
## 1. Estimated Salary 

df1$EstimatedSalary[df1$EstimatedSalary %in% boxplot.stats(df$EstimatedSalary)$out]
```
```{r}
## 2. Point Earned

df1$Point.Earned[df1$Point.Earned %in% boxplot.stats(df1$Point.Earned)$out]
```
```{r}
## 3. Credit Score

df1$CreditScore[df1$CreditScore %in% boxplot.stats(df1$CreditScore)$out]
```
```{r}
df2 <- df1[! df1$CreditScore %in% boxplot.stats(df1$CreditScore)$out, ]
```

```{r}
## Balance

df2$Balance[df2$Balance %in% boxplot.stats(df2$Balance)$out]
```
```{r}
data <- df2
```

## Developing the Model
## Spliting the data
```{r}
split <- sample.split(data, SplitRatio = 0.75)
split
```
```{r}
# defining the training and testing datasets

training <- subset(data, split == TRUE)
testing <- subset(data, split == FALSE)
```

## Building the model
```{r}
model <- glm(Exited~., training, family = "binomial")
summary(model)
```
## Checking the influence of the independent variables
```{r}
# 1. CreditScore
model1 <- glm(Exited~. -CreditScore, training, family = "binomial")
model1
```
Interpretation: Credit Score is a farily signifiant variable in the model. (Because AIC has decreased and Residual Deviance have increased after removing it)

```{r}
# Age
model1 <- glm(Exited~. -Age, training, family = "binomial")
model1
```
Interpretation: Age is a significant variable as both Residual Deviance and AIC have increased after removing it.

```{r}
# Tenure 
model1 <- glm(Exited~. -Tenure, training, family = "binomial")
model1
```
Interpretation: Tenure is not a significant variable due to the decrease in the AIC and Rsidual Deviance.

```{r}
# Balance 

model1 <- glm(Exited~.-Balance, training , family = "binomial")
model1
```
Interpretation: Balance is significant variable on the model as both Residual Deviance and AIC have increased.

```{r}
# NumOfProducts 

model1 <- glm(Exited~.-NumOfProducts, training , family = "binomial")
model1
```
# Interpretation: Number of Products is significant variable on the model as both Residual Deviance and AIC have increased.

```{r}
# HasCrCard

model1 <- glm(Exited~.-HasCrCard, training , family = "binomial")
model1
```
# Interpretation: Ownership of a Credit Card is not significant variable on the model as both Residual Deviance and AIC have decreased.

```{r}
# IsActiveMember

model1 <- glm(Exited~.-IsActiveMember, training , family = "binomial")
model1
```
Interpretation: Member Status is significant variable on the model as both Residual Deviance and AIC have increased.

```{r}
# EstimatedSalary

model1 <- glm(Exited~.-EstimatedSalary, training , family = "binomial")
model1
```
Interpretation: Estimated Salary is not significant variable on the model as both Residual Deviance and AIC have decreased.

```{r}
# Complain

model1 <- glm(Exited~.-Complain, training , family = "binomial")
model1
```
Interpretation: Complains are a significant variable on the model as both Residual Deviance and AIC have increased.

```{r}
# Satisfaction.Score

model1 <- glm(Exited~.-Satisfaction.Score, training , family = "binomial")
model1
```
Interpretation: Satisfaction Score is not significant variable on the model as both Residual Deviance and AIC have decreased.

```{r}
# Point.Earned

model1 <- glm(Exited~.-Point.Earned, training , family = "binomial")
model1
```
Interpretation: Points Earned are a significant variable on the model as both Residual Deviance and AIC have increased.

## Predicting the values for the model
```{r}
pred <- predict(model, testing, type = "response")
```

## Creating the Confusion Matrix to find the accuracy of the model
```{r}
table(Actual_Value = testing$Exited, Predicted_Value = pred>0.4)
table
```
## Finding the accuracy
```{r}
(1990 + 505) / (1990+1+1+505)
```
Interpreation: The model has an accuracy of 99%.

## Conclusion
The model has an accuracy of 99% making it a good predictor of customer churn in the bank.






