---
title: "Companies_EDA_Project"
author: "Alvin Kinoti"
date: "2023-11-20"
output: 
---

# 1000 COMPANIES EDA ANALYSIS REPORT

## Selecting the data 

```{r}
df <- read.csv("D:/Data Projects/Practice Files/1000_Companies.csv")
df
```
## Cleaning the data 
```{r}
sum(is.na.data.frame(df))
```
```{r}
df1 <- df[-4]
```

### Checking for outliers using a boxplot 

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)


ggplot(data = df1, aes(x= Marketing.Spend, y=Profit ))+
  geom_boxplot()+
  labs(x= "Marketing.Spend", y = "Profit")+
  ggtitle("Marketing Expenditure and Profit")
```


```{r}
df$Profit[df$Profit%in% boxplot.stats(df$Profit)$out]  
View(df)
```
### Removing the outliers

```{r}
df <- df[! df$Profit %in% boxplot.stats(df$Profit)$out, ]
df
```

## Data Visualization
### Histogram
#### Interpretation : the most profit is realized between 150000 and 200000.

```{r}
ggplot(data = df, aes(x = Profit))+
  geom_histogram(fill = "green", color = "black")+
  labs(x= "Profit", y = "Frequency")+
  ggtitle(" Profit Histogram")
```
### Line graph
#### Interpretation : there is a strong correlation between the Marketing Expenditure and Profit.

```{r}
ggplot(data = df, aes(x= Marketing.Spend, y = Profit))+
  geom_line(color = "red")+
  labs(x = "Marketing Spend", y = "Profit")+
  ggtitle("Marketing Spend vs Profit")
```

### Correlation
#### Marketing Expenditure and Profit
##### there is an almost perfect positive correlation between the Marketing Expenditure and Profit

```{r}
cor(df$Marketing.Spend, df$Profit)
```

#### R.D Spend and Profit
##### there is an almost perfect positive correlation between the R.D Spend and Profit

```{r}
cor.test(df$R.D.Spend, df$Profit)
```
#### Administration and Profit
##### There is a positive correlation between Administration and Profit

```{r}
cor.test(df$Administration, df$Profit)
```






