---
title: "Diabetes Prediction Model"
author: "Alvin Kinoti"
date: "2024-01-29"
output: word_document
---

## Load the libraries and the dataset
```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(caTools)

data <- read.csv("D:/Data Projects/Practice Files/Diabetes/diabetes.csv")
View(data)
```

## Dealing with any null values
```{r}
data <- na.omit(data)
```

##  Splitting the dataset into testing and training
```{r}
split <- sample.split(data, SplitRatio = 0.8)
split

training <- subset(data, split == TRUE)
testing <- subset(data, split == FALSE)
```
## Buidling the model
```{r}
model <- glm(Outcome~., training, family = "binomial")
summary(model)
```
## Checking the influence of non significant i..v on the model
```{r}
model <- glm(Outcome~.-SkinThickness, training, family = "binomial")
summary(model)
```
## SkinThickness is not a signifiant variable in the model. (Because AIC and Residual Devance have decreased after removing it)

```{r}
model <- glm(Outcome~.-Insulin, training, family = "binomial")
summary(model)
```
## Insulin is a signifiant variable in the model. (Because AIC and Residual Devance have increased after removing it)

```{r}
model <- glm(Outcome~.-Age, training, family = "binomial")
summary(model)
```
## Age is a signifiant variable in the model. (Because AIC and Residual Devance 
have increased after removing it)

## Predicting the values for the dataset

```{r}
res <- predict(model, testing, type = "response")
res
```
## Comparing the real values with the predicted values 
## threshold = 0.5, if probability =>0.5, patient is diabetic, 
## probability <0.5, patient is not diabetic
```{r}
testing
```
## Calculating the threshold
```{r}
res <- predict(model, training, type = "response")

library(ROCR)
ROCRPred <- prediction(res, training$Outcome)
ROCRPerf <- performance(ROCRPred, "tpr", "fpr")
```

## Plotting the threshold analysis graph
```{r}
plot(ROCRPerf, colorize=TRUE, print.cutoffs.at=seq(0.1, by=0.1))
```
## Computing the Confusion Matrix
```{r}
res <- predict(model, testing, type = "response")

table(Actual_Value = testing$Outcome, Predicted_Value = res>0.3)
table
```
## Calculating the Accuracy of the Model
```{r}
(88+42)/ (88+31+10+42)
```
## Conclusion: The Model has a 76% accuracy rate and is a good predictor.