---
title: "Restaurant Promotion Promotion Prediction Model"
author: "Alvin"
date: "2024-03-07"
output: word_document
---

## Loading the libraries and dataset
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(caTools)
```

```{r}
df <- read.csv("D:/Data Projects/Practice Files/Restaurants Revenue Prediction/Restaurant_revenue (1).csv")

View(df)
```

## Exploring the dataset
```{r}
summary(df)
```
```{r}
# Checking for any null values

sum(is.na(df))
```
```{r}
# Checking for any duplicates

sum(duplicated(df))
```
```{r}
## Dealing with outliers with the Monthly Revenue

df$Monthly_Revenue[ df$Monthly_Revenue %in% boxplot.stats(df$Monthly_Revenue)$out]
```
```{r}
restaurant <- df[! df$Monthly_Revenue %in% boxplot.stats(df$Monthly_Revenue)$out, ]

View(restaurant)
summary(restaurant)
```
## Visualizing the Dataset
```{r}
# Cusisine Types greater than the monthly revenue median with the number of customers

result1 <- restaurant %>% 
  filter(Monthly_Revenue > 270.513) %>% 
  group_by(Marketing_Spend) %>% 
  select(Cuisine_Type, Monthly_Revenue)
  
print(result1)
```
```{r}
# Cuisine Types with the highest demand

ggplot(data = restaurant, aes(x = Cuisine_Type, y = Number_of_Customers))+
  geom_col(fill = "maroon")+
  labs(x = "Cuisine Type", y = "Demand")+
  ggtitle("Demand per Cuisine Type")
```
```{r}
## Relationship between Demand and Cost of Meals

ggplot(data = restaurant, aes(x = Number_of_Customers, y = Menu_Price))+
  geom_point()+
  geom_smooth(method = "lm", se = F, color = "green")+
  labs(x = "Demand", y = "Cost of Meals")+
  ggtitle("Relationship between Demand and Cost of Meals")
```
```{r}
## The Relationship between Demand and Cost of Meals is almost Elastic.
```
## Designing the Promotions Prediction Model
```{r}
restaurant1 = restaurant[-c(4)]

split <- sample.split(restaurant1, SplitRatio = 0.75)
split
```

```{r}
training <- subset(restaurant1, split == TRUE)
testing <- subset(restaurant1 , split == FALSE)
```

```{r}
model1 <- glm(Promotions ~., training, family = "binomial")
summary(model1)
```
## Checking influence of Independent Variables on the Dependent Variable
```{r}
model <- glm(Promotions ~.-Number_of_Customers, training, family = "binomial")
model
```
## The Number of Customers (Demand) is significant to the model since Residual Deviance has Increased after removing it

```{r}
model <- glm(Promotions ~.-Menu_Price, training, family = "binomial")
model
```
## The Menu Price is insignificant to the model since AIC and Residual Deviance have deccreased after removing it.

```{r}
model <- glm(Promotions ~.-Marketing_Spend, training, family = "binomial")
model
```
## The Marketing Expenditure (Marketing_Spend) is insignificant to the model since AIC and Residual Deviance have deccreased after removing it.

```{r}
model <- glm(Promotions ~.-Average_Customer_Spending, training, family = "binomial")
model
```
## The Average_Customer_Spending) is significant to the model since AIC and Residual Deviance have increased after removing it.

```{r}
model <- glm(Promotions ~.-Reviews, training, family = "binomial")
model
```
## The Reviews is insignificant to the model since AIC and Residual Deviance have decreased after removing it.

```{r}
model <- glm(Promotions ~.-Monthly_Revenue, training, family = "binomial")
model
```
## The Monthly Revenue is insignificant to the model since AIC and Residual Deviance have decreased after removing it.

## Predicting the Values for the dataset
```{r}
res <- predict(model1, testing, type = "response")
res
```
## Creating the Confusion Matrix to find the accuracy of the model
```{r}
table(Actual_Value = testing$Promotions, Predicted_Value = res>0.5)
table
```
## Finding the Accuracy
```{r}
(97+45) / (97+34+108+45)
```




