---
title: "Restaurants Revenue EDA"
author: "Alvin"
date: "2024-08-10"
output: pdf_document
---

## Setting the working directory
```{r}
setwd("D:/Personal Projects 2/R Programming")

getwd()
```
## Loading the required libraries
```{r}
library(ggplot2)
library(corrplot)
library(tidyverse)
library(dplyr)
```

## Loading the dataset
```{r}
df <- read.csv("D:/Personal Projects 2/Practice Files July 2024/Restaurants Revenue Prediction/Restaurant_revenue (1).csv")
```

## ETL
```{r}
# the first 5 rows

head(df)
```
```{r}
## summary statistics of the dataset

summary(df)
```
From the above results, the shape of the dataset is (1000, 8).

```{r}
## checking for any null values 

sum(is.na(df))
```

```{r}
## checking for any duplicates

sum(duplicated(df))
```
There are no duplicates or null values in the dataset

N/B - Due to the small size of the dataset, the outliers are not removed. 

## Analyzing and visualizing the dataset
```{r}
## Identifying the total number of customers

cat("Total number of Customers:",sum(df$Number_of_Customers))
```
```{r}
## Average menu price

cat("Average Menu Price:", mean(df$Menu_Price))
```
```{r}
## Average Marketing Spend

cat("Average Marketing Spend:", mean(df$Marketing_Spend))
```
```{r}
## Average Monthly Revenue

cat("Average Monthly Revenue:",mean(df$Monthly_Revenue))
```
```{r}
## Median Reviews

cat("Median Reviews:", median(df$Reviews))
```
```{r}
## Most preferred cuisine

ggplot(data = df, aes(x = Cuisine_Type, y = Number_of_Customers))+
  geom_col(fill = "maroon")+
  labs( x = "Cuisine Type" , y = "Number of Customers")+
  ggtitle("Demand per Cuisine", subtitle = "(Analysis of Cusdtomer Preference For Each Cuisine)")
```
The Japanese is the most in demand Cuisine

```{r}
## Most profitable cuisine

ggplot(data = df, aes(x = Monthly_Revenue, y = Cuisine_Type))+
  geom_col(fill = "darkgreen")+
  labs(x = "Revenue", y = "Cuisine")+
  ggtitle("Revenue Per Cuisine")
```
The Japanese is the most profitable cuisine.

```{r}
## The relationship between numeric variables

numeric <- df[-c(4)] ##removing categorical data

correlation <- cor(numeric)

corrplot(correlation, method =  "number", order = "alphabet", title = "Relationship Between Variables")
```
Tip: The boxxes without numbers imply that the correlation between the two variables is zero.


```{r}
## t-test for the numerical values

library(readr)

t_test_price_revenue <- t.test(df$Monthly_Revenue, df$Menu_Price)
t_test_price_revenue
```
The null hypothesis is rejected since there exists a statistical difference between the two populations.

```{r}
t_test_reviews_marketing <- t.test(df$Reviews, df$Marketing_Spend)
t_test_reviews_marketing
```
The two means are significantly different as specified by the null hypothesis, thus the null hypothesis is discredited. 

```{r}
## Investigating the line of best fit

mae_profit = mean(abs(df$Monthly_Revenue - df$Menu_Price))
mae_campaign_effectiveness = mean(abs(df$Reviews - df$Marketing_Spend))

## printing the above two results
cat("Mean Absolute Error for Profit:", mae_profit, "\n")
cat("Mean Absolute Error for Campaign Effectiveness:", mae_campaign_effectiveness, "\n")
```
The mean absolute error for Campaign Effectiveness indicates a better line of best fit than the mean absoulte error for profit as it it more closer to zero.




