---
title: "Restaurant Revenue Prediction Project"
author: "Alvin"
date: "2025-01-25"
output: word_document
---

## Loading the dataset
```{r}
df <- read.csv("D:/others/Personal Projects 2/Practice Files July 2024/Restaurant Revenue Prediction Dataset/restaurant_data.csv")

head(df)
```
## Loading the required libraries
```{r}
library(dplyr)
library(tidyverse)
library(caTools)
library(corrplot)
```
## Exploring the dataset
```{r}
## shape of the dataset
dim(df)
```
```{r}
## summary statistics 

summary(df)
```
## Data Wrangling
```{r}
# checking for any null values

sum(is.na(df))
```
```{r}
# checking for any duplicates

sum(duplicated(df))
```
```{r}
# dealing with outliers

df1 <- df[! df$Revenue %in% boxplot.stats(df$Revenue)$out, ]
dim(df1)
```
```{r}
# outliers removed 

8368 - 8266
```
## Data Mining
## 1. KPIs
```{r}
# Average Revenue

data <- df1

mean(data$Revenue)
```
```{r}
# No of restaurants

length(data$Name)
```

```{r}
# total marketing budget

sum(data$Marketing.Budget)
```
```{r}
## average rating

mean(data$Rating)
```
```{r}
# Average Ambience Score 

mean(data$Ambience.Score)
```
## 1. Univariate Analysis
```{r}
## Revenue Distribution

ggplot(data = data, aes(x = Revenue))+
  geom_density(show.legend = "Location", color = "darkgrey", fill = "maroon")+
  labs(X= "Revenue") +
  ggtitle("Revenue Distribution")
```
```{r}
## ratings distribution

ggplot(data = data, aes(x = Rating))+
  geom_histogram(binwidth = 0.3, fill = "blue", color = "black")+
  labs(X= "Ratings")+
  ggtitle("Ratings Distribution")
```
## 2. Bivariate Analysis
```{r}
## Revenue per location

ggplot(data = data, aes(x =  Location, y = Revenue))+
  geom_col(fill = "darkgreen")+
  labs(x = "Locations", y = "Revenue")+
  ggtitle("Revenue Generated Per Location")
```
```{r}
## Social Media Followers per Location

ggplot(data = data, aes(y =  Location, x = Social.Media.Followers))+
  geom_col(fill = "darkgreen")+
  labs(y = "Locations", x = "Social Media Followers")+
  ggtitle("Revenue Generated Per Location")
```
## Developing the Prediction Model
```{r}
## splitting the data

numeric_data = data[-c(1,2,3,14)]

split <- sample.split(numeric_data, SplitRatio = 0.75)
split
```
```{r}
## dividing the data into training and testing datasets

training = subset(numeric_data, split == TRUE)
testing = subset(numeric_data, split == FALSE)
```

```{r}
## Implementing the model

model_lm <- lm(Revenue~., data =  training)
summary(model_lm)
```
## Conclusion:
1. The p-value is less than the confidence level thus we reject the null hypothesis and the model has a 95% accuracy therefore making it an excellent predictor of restaurant revenue. 
