---
title: "Store Sales EDA"
author: "Alvin Kinoti"
date: "2024-01-15"
output: word_document
---

## Loading the required libraries
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

## Loading the dataset
```{r}
library(readxl)
sales <- read_excel("D:/Data Projects/Practice Files/US Stores Sales/sales.xlsx")
View(sales)
```

## Cleaning the dataset
```{r}
summary(sales)
```

```{r}
sum(duplicated(sales))
```
### Dealing with Outliers
```{r}
sales$Profit[sales$Profit %in% boxplot.stats(sales$Profit)$out]
```

```{r}
sales <- sales[! sales$Profit %in% boxplot.stats(sales$Profit)$out, ]
sales                
```

```{r}
sales$Margin[sales$Margin %in% boxplot.stats(sales$Margin)$out]
```

```{r}
sales <- sales[! sales$Margin %in% boxplot.stats(sales$Margin)$out, ]
sales
```

```{r}
sales$Sales[sales$Sales %in% boxplot.stats(sales$Sales)$out]
```


```{r}
sales <- sales[! sales$Sales %in% boxplot.stats(sales$Sales)$out, ]
sales
```

```{r}
sales$COGS[sales$COGS %in% boxplot.stats(sales$COGS)$out]
```

```{r}
sales <- sales[! sales$COGS %in% boxplot.stats(sales$COGS)$out, ]
sales
```


```{r}
sales$`Total Expenses`[sales$`Total Expenses` %in% boxplot.stats(sales$`Total Expenses`)$out]
```


```{r}
sales <- sales[! sales$`Total Expenses` %in% boxplot.stats(sales$`Total Expenses`)$out, ]
sales
```


```{r}
sales$Marketing[sales$Marketing %in% boxplot.stats(sales$Marketing)$out]
```

```{r}
sales <- sales[!sales$Marketing %in% boxplot.stats(sales$Marketing)$out, ]
sales
```


```{r}
sales <- sales[! sales$Inventory %in% boxplot.stats(sales$Inventory)$out, ]
sales
```

```{r}
sales <- sales[! sales$`Budget Profit` %in% boxplot.stats(sales$`Budget Profit`)$out, ]
sales
```


```{r}
sales <- sales[! sales$`Budget COGS` %in% boxplot.stats(sales$`Budget COGS`)$out, ]
sales
```

```{r}
sales <- sales[!sales$`Budget Margin` %in% boxplot.stats(sales$`Budget Margin`)$out, ]
sales
```


```{r}
sales <- sales[!sales$`Budget Sales` %in% boxplot.stats(sales$`Budget Sales`)$out, ]
sales
```


```{r}
summary(sales)
```
## Visualizing and analyzing the dataset
### Profit per Consumer Market
```{r}
ggplot(data = sales, aes(x = Market, y = Profit))+
  geom_col(color = "blue")+
  labs(x = "Consumer Market", y = "Profit")+
  ggtitle("Profit per Consumer Market ")
  
```
### Relationship between Profit and Cost of Production

```{r}
ggplot(data = sales , aes(x= COGS, y = Profit))+
  geom_line()+
  labs(x= "Cost of Production", y = "Profit")+
  ggtitle("Relationship between Profit and Cost of Production")
```
### Highest Accounted Profit Recorded
```{r}
ggplot(data = sales, aes(x=`Budget Profit`))+
  geom_histogram(bins = 10)+
  labs(x = "Accounted Profit", y = "Frequency")+
  ggtitle("Accounted Profit Frequency")
```
## Product Type With The Highest Accounted Profit
```{r}
ggplot(data = sales, aes(x = "", fill = `Product Type`, y = `Budget Profit`))+
  geom_col()+
  coord_polar(theta = "y")+
  ggtitle("Accounted Profit per Product Type")
```
## State with the Highest Projected Sales
```{r}
ggplot(data = sales, aes(y= State, x = `Budget Sales`))+
  geom_col(fill = "black")+
  labs(y = "State", x = "Projected Sales")+
  ggtitle("State with the Highest Projected Sales")
```
### Checking for correlations between variables

#### Profit and Accounted Profit
```{r}
cor.test(sales$Profit, sales$`Budget Profit`)
```
#### Budget Sales and Marketing
```{r}
cor.test(sales$`Budget Sales`, sales$Marketing)
```
#### Budget Margin and Total Expenses 
```{r}
cor.test(sales$`Budget Margin` , sales$`Total Expenses`)
```

#### Budgeted Production Cost and Budget Sales
```{r}
cor.test(sales$`Budget COGS`, sales$`Budget Sales`)
```

