---
title: "Student Exam Score EDA & Prediction Model"
author: "Alvin"
date: "2024-11-15"
output:
---

## Loading the Dataset
```{r}

df <- read.csv("D:/others/Personal Projects 2/Practice Files July 2024/Student Performance Factors/StudentPerformanceFactors.csv")
```

## Loading the required libraries
```{r}
library(dplyr)
library(tidyverse)
library(caTools)
library(corrplot)
```

## Exploring the dataset
```{r}

# summary statistics

summary(df)
```
## Bottom 4 Rows
```{r}
tail(df)
```

## Wrangling the Dataset
```{r}
## shape of the dataset

dim(df)
```

## Removing Duplicates
```{r}

df <- unique(df)

dim(df)
```
## Removing Rows with Missing Values
```{r}
dfa <- na.omit(df) # Delete rows with missing values 
dfa
```

## Checking For Any Outliers
```{r}
## hours studies

df1 <- dfa[! df$Hours_Studied %in% boxplot.stats(dfa$Hours_Studied)$out, ]  
df1
```
```{r}
## Attendance Column

df2 <- df1[! df$Attendance %in% boxplot.stats(df$Attendance)$out, ] 
df2
```
```{r}
#exam score column

df3 <- df2[! df2$Exam_Score %in% boxplot.stats(df$Exam_Score)$out, ] 
df3

data <- df3
```

## Exploratory Data Analysis

## 1. Univariate Analysis
```{r}
## Frequency of Hours Studied

library(ggplot2)

ggplot(data = data, aes(x = Hours_Studied))+
  geom_histogram(binwidth = 0.2, fill = "blue", color = "black")+
  labs(X = "Hours Studied", y= "Count")+
  ggtitle("Frequency of Hours Studied")
```
### The Hours Studied has a normal distribution.

```{r}
## distribution of exam scores

ggplot(data = data, aes(x = Exam_Score))+
  geom_histogram(binwidth = 0.2, fill = "blue", color = "black")+
  labs(X = "Exam Score", y= "Count")+
  ggtitle("Exam Score Distribution")
```
### The Exam Score has a normal distribution.

## 2. Bivariate analysis
```{r}
## Relationship Between Parental Involvement and Exam Score

clean_data <- data[!is.na(data$Parental_Involvement), ] ## removing undetected null values in the parental involvement column

ggplot(data = clean_data, aes(x = Exam_Score, y= Parental_Involvement))+
  geom_col(fill = "maroon")+
  labs(x = "Exam Scores", y = "Parental Involvement Level")+
  ggtitle("Correlation Between Exam Scores and Levels of Parental Involvement", 
          subtitle = "A simple Psycho-Analysis affecting student performance based on parental involvement")
```
```{r}
## relationship between school type and exam score

ggplot(data = clean_data, aes(x = School_Type, y = Exam_Score))+
  geom_col(color = "darkblue")+
  labs(x = "School Type", y = "Exam Score")+
  ggtitle("Exam Scores For Each Score", 
          subtitle = "An investigation to which school has the highest performers")
```
### Colcustion: Public Schools have more high performers than private schools.

## Developing the Exam Score Prediction model
```{r} 
## splitting the data

numeric_data = clean_data[c(1,2,6,7,10,15,20)]

split <- sample.split(numeric_data, SplitRatio = 0.8)
split
```
```{r}
## defining the training and testing datasets

training = subset(numeric_data, split == TRUE)
testing = subset(numeric_data, split == FALSE)
```


```{r}
## implementing the model

model_lm <- lm(Exam_Score~., data = training)
summary(model_lm)
```
## Interpretation: 

##### From the above output, only the sleep hours don't affect exam scores significantly. The other dependent variables have a significant impact on the exam score.

##### The null hypothesis is rejected since the p-value is less than the degrees of freedom, thus the dependent varibales do predict the exam score effectively.

##Conclusion: The model has an accuracy of 80% thus making it a good predictor of the students' exam score.


## Developing the R shiny App for the above model and analysis
```{r}
## loading the R shiny app library

library(shiny)
library(caret)
```

## Desigining the R shiny app
```{r}
# Define UI for the Shiny app

ui <- fluidPage(
  titlePanel("Student Exam Score Analysis and Prediction"),
  
  sidebarLayout(
    sidebarPanel(
      fileInput("file", "Upload CSV File", accept = ".csv"),
      selectInput("x_var", "Select X Variable:", choices = NULL),
      selectInput("y_var", "Select Y Variable:", choices = NULL),
      selectInput("chart_type", "Choose Chart Type:", 
                  choices = c("Histogram", "Boxplot", "Scatterplot", "Correlation Matrix")),
      actionButton("analyze", "Generate Chart"),
      hr(),
      h3("Prediction Section"),
      numericInput("hours_studied", "Hours Studied:", value = 5, min = 0, max = 24),
      numericInput("attendance", "Attendance (%):", value = 80, min = 0, max = 100),
      actionButton("predict", "Predict Exam Score")
    ),
    
    mainPanel(
      tabsetPanel(
        tabPanel("Visualization", plotOutput("plot")),
        tabPanel("Summary", verbatimTextOutput("summary")),
        tabPanel("Prediction", verbatimTextOutput("prediction"))
      )
    )
  )
)

# Define the server logic
server <- function(input, output, session) {
  data <- reactiveVal(NULL)
  
  observeEvent(input$file, {
    req(input$file)
    df <- read.csv(input$file$datapath)
    data(df)
    
    updateSelectInput(session, "x_var", choices = names(df))
    updateSelectInput(session, "y_var", choices = names(df))
  })
  
  output$summary <- renderPrint({
    req(data())
    summary(data())
  })
  
  output$plot <- renderPlot({
    req(data())
    df <- data()
    
    if (input$chart_type == "Histogram") {
      ggplot(df, aes_string(input$x_var)) + 
        geom_histogram(binwidth = 10, fill = "blue", color = "white") +
        theme_minimal()
    } else if (input$chart_type == "Boxplot") {
      ggplot(df, aes_string(y = input$x_var)) + 
        geom_boxplot(fill = "blue", color = "black") +
        theme_minimal()
    } else if (input$chart_type == "Scatterplot") {
      req(input$y_var)
      ggplot(df, aes_string(x = input$x_var, y = input$y_var)) + 
        geom_point(color = "blue", size = 3) +
        theme_minimal()
    } else if (input$chart_type == "Correlation Matrix") {
      corrplot(cor(df[sapply(df, is.numeric)]), method = "circle")
    }
  })
  
  output$prediction <- renderPrint({
    req(data())
    df <- data()
    
    model <- train(Exam_Score ~ Hours_Studied + Attendance, data = df, method = "lm")
    prediction <- predict(model, 
                          newdata = data.frame(Hours_Studied = input$hours_studied, 
                                               Attendance = input$attendance))
    paste("Predicted Exam Score:", round(prediction, 2))
  })
}
```


## Running the R shiny Application
```{r}
# Run the application 

shinyApp(ui = ui, server = server)
```

