---
title: "Video Game Sales EDA Project"
author: "Alvin"
date: "2024-07-13"
output: word_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

The Main purpose of this EDA Project is to conduct a simple but robuts Exploratory Data Analysis exericise on a video games dataset.

It involves:
1. Loading the dataset
2. Exploring the dataset
3. Cleaning the dataset (removing null values, cleaning, detecting and removing outliers)
4. Visualizng the dataset
5. Exploratory Data Analysis
6. Conclusion

## Setting the working directory
```{r}
setwd("C:/Users/Alvin/Desktop/Data Projects/R Programming")
getwd()
```
## Loading the dataset
```{r}
df <- read.csv("C:/Users/Alvin/Downloads/Compressed/Video Games Sales Dataset/Video_Games_Sales_as_at_22_Dec_2016.csv")
```

## Investigating the dataset

## checking a brief overview of the dataset
```{r}
summary(df)
```
The above results show the various metrics of the columns within the dataset.

## Checking the shape of the dataset
```{r}
dim(df)
```
The dataset has 16719 rows and 16 columns

## Cleaning the dataset
### checking the number of empty values in the dataset
```{r}
sum(is.na(df))
```
### Removing the empty values 
```{r}
df1 <- na.omit(df)

# checking for any remaining empty values
sum(is.na(df1))
```
### Checking for any duplicates
```{r}
sum(duplicated(df1))
```
There is no duplicated values in the cleaned dataset.

## Checking for any outliers in the dataset
### Checking in the NA_Sales Column
```{r}
df1$NA_Sales[df1$NA_Sales %in% boxplot.stats(df1$NA_Sales)$out]
```
### Removing the outliers
```{r}
df2 <- df1[! df1$NA_Sales %in% boxplot.stats(df1$NA_Sales)$out, ]
```

### Removing outliers in Other Columns
### 1. EU_Sales
```{r}
df3 <- df2[! df2$EU_Sales %in% boxplot.stats(df2$EU_Sales)$out, ]
```

### 2. JP_Sales
```{r}
df4 <- df3[! df3$JP_Sales %in% boxplot.stats(df3$JP_Sales)$out, ]
```

### 3. Other_Sales 
```{r}
df5 <- df4[! df4$Other_Sales %in% boxplot.stats(df4$Other_Sales)$out, ]
```

### 4. Global_Sales
```{r}
df6 <- df5[! df5$Global_Sales %in% boxplot.stats(df5$Global_Sales)$out, ]
```

### 5. Critic_Score
```{r}
df7 <- df6[! df6$Critic_Score %in% boxplot.stats(df6$Critic_Score)$out, ]
```

### 6. Critic_Count
```{r}
df8 <- df7[! df7$Critic_Count %in% boxplot.stats(df7$Critic_Count)$out, ]
```

### 7.User_Score
```{r}
df9 <- df8[! df8$User_Score %in% boxplot.stats(df8$User_Score)$out, ]
```

### 8. User_Count
```{r}
df10 <- df9[! df9$User_Score %in% boxplot.stats(df9$User_Score)$out, ]

data <- df10
```

## Exploratory data analysis

### Visualizing the dataset

### 1. Plotting the Developer  with the highest Demand

```{r}
#loading the required libraries
library(ggplot2)
library(dplyr)
library(tidyverse)

ggplot(data = data, aes(x=Genre ))+
  geom_bar( fill = "skyblue")+
  labs(x= "Genre", y = "Demand")+
  ggtitle("Demand Per Genre")+
  theme(plot.title = element_text(hjust=0.5))
```
### 2. Comparison Between Other Sales and Global Sales
```{r}
ggplot(data = data, aes(x = Other_Sales, y = Global_Sales))+
  geom_line(stat = "identity", position = "identity", color = "darkgreen")+
  labs(x= "Other Sales", y = "Global Sales")+
  ggtitle("Comparison Between Other Sales and Global Sales")+
  theme(plot.title = element_text(hjust = 0.5))
```
### 3. plotting the correlation plot between the variables
```{r}
numeric = data[-c(1:5,8,15:16)]
 
correlation = cor(numeric)

library(corrplot)
corrplot(correlation, method = 'color', order = 'alphabet')
```
### 4. Finding the paired t-test for global and other sales
```{r}
library(readr)

t_test_sales <- t.test(data$Global_Sales, data$Other_Sales, paired = TRUE)
t_test_sales
```
### 5. Finding the paired t-test for EU and NA sales
```{r}
library(readr)

t_test_sales_2 <- t.test(data$NA_Sales, data$EU_Sales, paired = TRUE)
t_test_sales_2
```
### 6. MeaN Absolute Errors (MAE) for the above two pairs
```{r}
mae_global_other_sales = mean(abs(data$Global_Sales - data$Other_Sales))
mae_na_eu_Sales = mean(abs(data$NA_Sales - data$EU_Sales))

## printing the above two results
cat("Mean Absolute Error for Global and Other Sales:", mae_global_other_sales, "\n")
cat("Mean Absolute Error for Na and Eu Sales:", mae_na_eu_Sales, "\n")
```
## Interpreation of the above two:

The Mean Absolute Error for Na and Eu Sales indicated a better fit of the data compared to the Mean Absolute Error for Global and Other Sales as it is much closer to zero, thus having a better line of best fit and a prediction model with a higher accuracy.


## Conclusion
1. Action Video Games lead with the highest demand of more than 750 while Puzzle games trail with a demand of less than 125.
2. The trend between Global Sales and Other Sales is on an upward trend.
3. There is a strong relationship between Global Sales, EU Sales, Other Sales and NA Sales. This could be attributed to a variety of factors e.g access to gaming devices, technological adfvancement in those markets etc.

## Recommendations:
1. There is need for more technological upgrades and marketing strategies in other markets aprt from the Global Sales, EU Sales, Other Sales and NA Sales markets.
2. More games should be produced and sold in the Sales, EU Sales, Other Sales and NA Sales markets to maximize on sales. 
3. More action games should be produced to boost revenue.
