---
title: "Alvin Week 3 Assignment"
author: "Alvin Kinoti"
date: "2023-11-17"
output: github_document
---

 
# 1. Operators:
## a. Create a vector `x` with values 1 to 10. Generate a new vector `y` where each element is the square of the corresponding element in `x` minus twice the element itself.

```{r}
x <- 1:10

y <- x^2 - 2*x
cat("Vector x:", x, "\n")
cat("Vector y:", y, "\n")

```
b. Given a vector `grades` representing exam scores, create a new vector `pass` that indicates whether each student passed (scored 50 or above) or failed.

```{r}
grades <- c(10, 50, 30 , 85, 50, 80, 40, 65, 55, 90)

pass <- ifelse(grades >= 50, "Pass", "Fail")

cat("Grades:", grades, "\n")
cat("Pass:", pass, "\n")
```

# 2. ggplot:
## a. Utilize the `mpg` dataset in ggplot to create a scatter plot comparing highway mileage (`hwy`) and city mileage (`cty`). Add color differentiation to the points in the scatter plot based on the class of the vehicle.

```{r}
library(ggplot2)
library(dbplyr)
library(tidyverse)
attach(mpg)
View(mpg)

ggplot(data = mpg, aes(x = hwy, y = cty))+
  geom_point(color = "red") +
  labs(x = "Highway Mileage", y = "City Mileage")+
  ggtitle("Highway Mileage vs City Mileage")
```
 ## b. Utilize the `iris` dataset to create a faceted plot comparing the distributions of sepal length (`Sepal.Length`) for each species. Ensure the plots have proper labels and a main title.
 
```{r}
attach(iris)
View(iris)

ggplot(iris, aes(x = Sepal.Length)) +
  geom_density(fill = "brown", color = "green") +
  facet_wrap(~Species, scales = "free_y") +
  labs(
    title = "Distribution of Sepal Length by Species",
    x = "Sepal Length",
    y = "Density",
    caption = "Source: Iris dataset"
  )
```
 
```
# 3. Loops
## a. Write a function to generate the first 10 numbers in the Fibonacci sequence. Use a loop to print each number

```{r}
fibonacci_sequence <- function(n) {
  fib <- numeric(n)
  fib[1] <- 0
  fib[2] <- 1
  
  for (i in 3:n) {
    fib[i] <- fib[i - 1] + fib[i - 2]
  }
  
  return(fib)
}


for (num in fibonacci_sequence(10)) {
  cat(num, " ")
}

```


# b. Create two matrices `A` (2x3) and `B` (3x2) filled with random values. Write a loop to multiply these matrices and store the result in a new matrix `C`.

```{r}

set.seed(456)

X <- matrix(runif(3 * 4), nrow = 3, ncol = 4)


Y <- matrix(runif(4 * 2), nrow = 4, ncol = 2)


Z <- matrix(0, nrow = nrow(X), ncol = ncol(Y))

for (i in 1:nrow(X)) {
  for (j in 1:ncol(Y)) {
    for (k in 1:ncol(X)) {
      Z[i, j] <- Z[i, j] + X[i, k] * Y[k, j]
    }
  }
}


cat("Matrix X:\n", X, "\n")
cat("Matrix Y:\n", Y, "\n")
cat("Matrix Z (Result of X * Y):\n", Z, "\n")

```

# 4. Functions:
## a. Write a recursive function to calculate the factorial of a given positive integer.

```{r}
factorial <- function(n) {
  if (n == 0 || n == 1) {
    return(1)
  } else {
    return(n * factorial(n - 1))
  }
}

n <- 6
result <- factorial(n)

cat("The factorial of", n, "is", result, "\n")
```

## b. Develop a function that takes an integer as input and returns a Boolean indicating whether it is a prime number or not.

```{r}
is_prime <- function(x) {
  if (x <= 1) {
    return(FALSE)  
  }
  
  for (i in 2:sqrt(x)) {
    if (x %% i == 0) {
      return(FALSE)  
    }
  }
  
  return(TRUE)  
}

# For Instance
num <- 23
result <- is_prime(num)

cat("Is", num, "a prime number?", result, "\n")
```


# 5. R Markdown:
## a. Create an R Markdown document that loads a dataset of your choice, performs some exploratory data analysis, and generates visualizations using ggplot. 

### Loading the dataset
```{r}


df <- read.csv("D:/Data Projects/Bounty Analytica/R/1000_Companies.csv")
View(df)

summary(df)
tail(df)
head(df)

```

### Detecting the oultiers in the Profit Column

```{r}
df$Profit[df$Profit %in% boxplot.stats(df$Profit)$out]
View(df)
```

### Removing the outliers in the Profit Column

```{r}
df <- df[! df$Profit %in% boxplot.stats(df$Profit)$out, ]
df
```

## Visualizing the data
### Histogram

```{r}
ggplot(data = df, aes(x = Profit)) +
geom_histogram(fill= "green", color = "black") +
labs(x = "Profit", y = "Frequency") +
ggtitle("Profit histogram")

```

### Line graph 

```{r}
ggplot(data = df, aes(x=Marketing.Spend, y = Profit))+
geom_line(color = "red")+
labs(x= "Marketing Expenditure", y = "Profit")+
ggtitle("Relationship between Marketing Expenditure and Profit")
```

### Scatterplot 

```{r}
ggplot(data = df, aes(x=Administration, y = Profit))+
geom_point() +
geom_smooth(method = "lm", se = F, color = "brown") +
labs(x = "Administration", y = "Profit") +
ggtitle("Administration vs Profit")
```

## Correlation tests
### Marketing Expenditure and Profit
```{r}
cor.test(df$Marketing.Spend, df$Profit)
```

### Adminisration and Profit

```{r}
cor.test(df$Administration, df$Profit)
```

### R.D.Spend and Profit

```{r}
cor.test(df$R.D.Spend, df$Profit)
```

